---
title: "Halton Sequences for Randomized Quasi-Monte Carlo (RQMC)"
output: rmarkdown::html_vignette
vignette: >
  %\VignetteIndexEntry{halton}
  %\VignetteEngine{knitr::rmarkdown}
  %\VignetteEncoding{UTF-8}
---

```{r, include = FALSE}
knitr::opts_chunk$set(
  collapse = TRUE,
  comment = "#>"
)
```

The original algorithm was written by [Art B. Owen](http://statweb.stanford.edu/~owen/) and can be found [here](http://statweb.stanford.edu/~owen/code/). I translated the internal function `randradinv`, which produces randomized (scrambled) van der Corput sequences into **C**, and it is where most of the computation is done.

The code snippet below shows how to use the `rhalton` function and is taken from Figure 5 of the paper found at the above link. We compare the standard error of the RQMC estimate to the CMC estimate; unsurprisingly RQMC is several of orders less variable.

```{r setup}
library(mcfun)

f <- function(v){sum(v)^2} # Example function that is easy for RMQC

# RQMC
R <- 10
n <- 5000
p <- 20
stride <- 1000 # Or replace 1000 by nthprime(0,getlength=TRUE)
muvec <- rep(0,R)
for(r in 1:R){
    x <- rhalton(n,p,singleseed = r*stride)
    muvec[r] <- mean( apply( x,1,f ) )
}
mu <- mean(muvec)
se <- sqrt(var(muvec)/R)

# CMC (crude Monte Carlo)
muvec_cmc <- rep(0,R)
for(r in 1:R){
  x <- matrix(data = runif(n*p),nrow = n,ncol = p)
  muvec_cmc[r] <- mean( apply( x,1,f ) )
}
mu_cmc <- mean(muvec_cmc)
se_cmc <- sqrt(var(muvec_cmc)/R)

knitr::kable(matrix(data=c(mu,se,mu_cmc,se_cmc),nrow=2,ncol=2,byrow=T,dimnames=list(c("RQMC","CMC"),c("mean","se"))),caption = "Comparison of RQMC to crude Monte Carlo on a simple problem")
```
